---
title: "Naive Bayes Assignment - Fake News Classification"
author: 
- Audrya Kerenhappukh - Author 
- Ryan Kokke - Reviewer 

date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
  html_notebook:
    toc: yes
    toc_depth: 2
---


```{r message=TRUE, warning = TRUE, include = FALSE}
library(tidyverse)
library(dplyr)
library(wordcloud)
library(tm)
library(caret)
library(e1071)
library(readr)
```

## Business Understanding 
Develop a machine learning program to identify when an article might be fake news

## Import Document and Data Understanding 
```{r message=TRUE, warning = TRUE, include = FALSE}

url <- "https://raw.githubusercontent.com/HAN-M3DM-Data-Mining/assignments/master/datasets/NB-fakenews.csv"

rawDF <- read_csv(url) %>% na.omit # Assign to new variable and elimate NaN values


```
Because text could be really long we're trying to reduce the file size by making a new variable called 'author.title' 
```{r}
rawDF$author.title <- paste(rawDF$author, rawDF$title)

cbind(colnames(rawDF))
```

```{r}
# Create a new data set with only two variables: 'label' and 'author.title'
cleanDF <- rawDF[c(5,6)]

```


The labels in the dataset are still numerical, we would want to factor them into "unreliable" and "reliable"

```{r}

cleanDF$label <- factor(cleanDF$label, levels = c(1,0), labels = c("unreliable", "reliable")) %>% relevel("reliable")

```

Visually inspect the data by creating worldclouds
```{r message=TRUE, warning = FALSE, include = TRUE}
unreliable <- cleanDF %>% 
  filter(label == "unreliable")
reliable <- cleanDF %>% 
  filter(label == "reliable")

wordcloud(unreliable$author.title, max.words = 30, scale = c(4, 0.8), colors= c("indianred1","indianred2","indianred3","indianred"))
wordcloud(reliable$author.title, max.words = 30, scale = c(4, 0.8), colors= c("lightsteelblue1","lightsteelblue2","lightsteelblue3","lightsteelblue"))
```


## Preparation 
### Create corpus 

First we need to create a corpus, which refers to a collection of text documents. In our case each  is considered a text document

#### Creating corpus for Training data 

```{r}

rawCorpus <- Corpus(VectorSource(cleanDF$author.title))

```

```{r warning=FALSE}
cleanCorpus <- rawCorpus %>% tm_map(tolower) %>% tm_map(removeNumbers)
cleanCorpus <- cleanCorpus %>% tm_map(tolower) %>% tm_map(removeWords, stopwords()) %>% tm_map(removePunctuation)
```

```{r warning=FALSE}
cleanCorpus <- cleanCorpus %>% tm_map(stripWhitespace)

```
See the difference between Raw vs Clean Corpus 
```{r}
tibble(Raw = rawCorpus$content[1:2], Clean = cleanCorpus$content[1:2])
```

```{r}
cleanDTM <- cleanCorpus %>% DocumentTermMatrix
inspect(cleanDTM)
```
```{r}

```

Before we can start modelling, it is important to split the data sets into train and test sets. We are using 90% of datasets for training, and 10% for test.

```{r}
set.seed(1234)
trainIndex <- createDataPartition(cleanDF$label, p = .9, 
                                  list = FALSE, 
                                  times = 1)
head(trainIndex)

# Apply split indices to DF
trainDF <- cleanDF[trainIndex, ]
testDF <- cleanDF[-trainIndex, ]

# Apply split indices to Corpus
trainCorpus <- cleanCorpus[trainIndex]
testCorpus <- cleanCorpus[-trainIndex]

# Apply split indices to DTM
trainDTM <- cleanDTM[trainIndex, ]
testDTM <- cleanDTM[-trainIndex, ]

```

### Eliminating Words with lower frequencies 

```{r}
freqwords <- cleanDTM %>% findFreqTerms(1000)
trainDTM <- DocumentTermMatrix(trainCorpus, list(dictionary = freqwords))
testDTM <- DocumentTermMatrix(testCorpus, list(dictionary = freqwords))

inspect(trainDTM)
inspect(testDTM)

```

Another issue is that the Naive Bayes classifier is typically trained on categorical features. We now have numerical matrix with word counts. We will transform the counts into a factor that simply indicates whether the word appears in the document or not. We’ll first build our own function for this and then apply it to each column in the DTM.
```{r}
convert_counts <- function(x) {
  x <- ifelse(x > 0, 1, 0) %>% 
    factor(levels = c(0,1), labels = c("No", "Yes")) 
}

nColsDTM <- dim(trainDTM)[2]
trainDTM <- apply(trainDTM, MARGIN = 2, convert_counts)
testDTM <- apply(testDTM, MARGIN = 2, convert_counts)

head(trainDTM)
```
## Modeling and Evaluation

Create the model 
```{r}
nbayesModel <-  naiveBayes(trainDTM, trainDF$label, laplace = 1)
```

```{r}
predVec <- predict(nbayesModel, testDTM)
```

```{r}
cm <- confusionMatrix(predVec, testDF$label, positive = "unreliable", dnn = c("Prediction", "True"))

cm
```

## Visualization of Confusion Matrix 
```{r Chart Visualization }
table <- as.data.frame(cm$table)

plotTable <- table %>%
  mutate(goodbad = ifelse(table$Prediction == table$True, "good", "bad")) %>%
  group_by(True) %>%
  mutate(prop = Freq/sum(Freq))

ggplot(data = plotTable, mapping = aes(x = True, y = Prediction, fill = goodbad, alpha = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
  scale_fill_manual(values = c(good = "green", bad = "red")) +
  theme_bw() +
  xlim(rev(levels(table$True))) +
  theme(axis.text.y = element_text(angle=90, hjust=0.5))


```
